---
title: "2018 Trade Up Analysis"
author: "Richard Harries"
date: '`r format(Sys.Date(), "%A, %d %B %Y")`'
output:
  prettydoc::html_pretty:
    highlight: github
    theme: architect
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
source("Common Tools.R")
library(broom)
library(knitr)
library(kableExtra)
library(plotly)
library(readxl)
```

```{r data import, include=FALSE}
CBTU <-
  read_excel("CBTU PtC - Roll Out Year Q4 100419.xlsx",
    sheet = "Clean Data",
    skip = 0
  ) %>%
  mutate(
    Scheme = case_when(
      Scheme == "C" ~ "Control",
      Scheme == "MT" ~ "Match Trading"
    )
  )

# Calculate quarterly trading ratios and annual cumulative ratios
# for the baseline and programme years

for (i in c("Baseline ", "")) {
  for (j in 1:4) {
    CBTU[paste0("Trading Ratio ", i, "Q", j)] <- 100 * CBTU[paste0("Trading Income ", i, "Q", j)] /
      (CBTU[paste0("Trading Income ", i, "Q", j)] + CBTU[paste0("Non-trading Income ", i, "Q", j)])
  }
  CBTU[paste0("Trading Ratio ", i, "Cumulative")] <- 100 *
    rowSums(CBTU[grepl(paste0("Trading Income ", i, "Q"), names(CBTU))], na.rm = FALSE) /
    (rowSums(CBTU[grepl(paste0("Trading Income ", i, "Q"), names(CBTU))], na.rm = FALSE) +
      rowSums(CBTU[grepl(paste0("Non-trading Income ", i, "Q"), names(CBTU))], na.rm = FALSE))
}

# Calculate the change in quarterly and annual ratios
# between the two years

for (j in 1:4) {
  CBTU[paste0("Trading Ratio Change Q", j)] <-
    (CBTU[paste0("Trading Ratio Q", j)] - CBTU[paste0("Trading Ratio Baseline Q", j)])
  CBTU[paste0("Eligible To Claim Ratio Q", j)] <-
    (CBTU[paste0("Eligible To Claim Q", j)] / (CBTU[paste0("Trading Income Q", j)] + CBTU[paste0("Non-trading Income Q", j)])) * 100
}

CBTU["Trading Ratio Change Cumulative"] <-
  CBTU["Trading Ratio Cumulative"] -
  CBTU["Trading Ratio Baseline Cumulative"]
```

## Exploratory Data Analysis

There are 99 organisations in the sample, with up to 20 data points recorded for each (trading income and non-trading income, for each quarter, in the baseline year and programme year, plus four quarterly Trade Up grant awards). However, the School for Social Entrepreneurs has only provided 1,822 data points (92%) out of a potential 1,980.

**Trading ratios**

*Performance*

Mean trading ratios fell for organisations in the Control Group and rose for those in the Match Trading Group. This is consistent with the core hypothesis of Trade Up.

```{r means, echo = FALSE}
CBTU %>%
  group_by(Scheme) %>%
  summarise(
    n = n(),
    mean_BASELINE_C = mean(`Trading Ratio Baseline Cumulative`, na.rm = TRUE),
    mean_TRADEUP_C = mean(`Trading Ratio Cumulative`, na.rm = TRUE),
    mean_Change_C = mean(`Trading Ratio Change Cumulative`, na.rm = TRUE)
  ) %>%
  kable(
    col.names = c(
      "Group","n",
      "Baseline year", "Trade Up year", "Change"
    ),
    digits = 1, format = "html"
  ) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

The same pattern occurs for the median trading ratios although the difference is less marked, reflecting the fact that medians are less influenced by outliers (see boxplot below).

```{r medians, echo = FALSE}
CBTU %>%
  group_by(Scheme) %>%
  summarise(
    n = n(),
    median_BASELINE_C = median(`Trading Ratio Baseline Cumulative`, na.rm = TRUE),
    median_TRADEUP_C = median(`Trading Ratio Cumulative`, na.rm = TRUE),
    median_Change_C = median(`Trading Ratio Change Cumulative`, na.rm = TRUE)
  ) %>%
  kable(
    col.names = c(
      "Group", "n",
      "Baseline year", "Trade Up year", "Change"
    ),
    digits = 1, format = "html"
  ) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```
*Significance*

Tests for significance can be unreliable for small samples. Two tests that are appropriate for the Trade Up dataset are the Kruskal-Wallis Test and the Mann-Whitney-Wilcoxon Test. Neither suggests there is a significant difference between the Control Group and the Match Trading Group. This is **not** consistent with the core hypothesis of Trade Up.

```{r tests, echo = FALSE}
CBTU %>%
  mutate(Group = as.factor(Scheme),
         `Trading Ratio Change` = `Trading Ratio Change Cumulative`) %>%
  kruskal.test(`Trading Ratio Change` ~ Group, .)
CBTU %>%
  mutate(Group = as.factor(Scheme),
         `Trading Ratio Change` = `Trading Ratio Change Cumulative`) %>%
  wilcox.test(`Trading Ratio Change` ~ Group, .)
```

## Plots

**Distributions**

Four community businesses in the Match Trading Group saw their trading ratios increase by more than 50% in 2018, compared to maximum growth of 40% in the Control Group. By contrast, the largest trading ratio reduction occured in the Match Trading Group, falling by more than 50%. 

```{r distributions, echo=FALSE, warning=FALSE}
p <-
  CBTU %>%
    rename("Trading ratio" = "Trading Ratio Change Cumulative") %>%
    ggplot(aes(Scheme, `Trading ratio`,
      text = paste0(round(`Trading ratio`, 2),"\n", Sector)
    )) +
    geom_boxplot() +
    geom_point(aes(colour = `Sector`)) +
    labs(
      title = "Boxplot: Increase/decrease in trading ratios",
      x = "", y = "Percentage point change"
    ) +
  theme(legend.position = "none")
ggplot_ptc(p, "text")

p <-
  CBTU %>%
    ggplot() +
    geom_histogram(aes(
      x = `Trading Ratio Change Cumulative`,
      colour = 24, fill = Sector),
    bins = 5
    ) +
    guides(colour = FALSE) +
    labs(
      title = "Histogram: Increase/decrease in trading ratios",
      x = "Percentage point change", y = ""
    ) +
    facet_wrap(vars(Scheme))
ggplot_ptc(p, "y")

p <-
  CBTU %>%
    ggplot() +
    geom_histogram(aes(
      x = `Trading Ratio Change Cumulative`,
      colour = 24, fill = Sector
    ),
    bins = 10, show.legend = FALSE
    ) +
    labs(
      title = "Histogram: Increase/decrease in trading ratios, by group and location",
      x = "Percentage point change", y = ""
    ) +
    facet_wrap(vars(Scheme, Location), ncol = 5)
ggplot_ptc(p, "y")
```

**Trends**

Trading ratios went up in both the Control Group and the Match Trading Group across 2018, ending up just below 80% in both cases. However, compared to the baseline year, quarterly trading ratios generally **fell** in both groups, with the Control Group doing relatively **better** than the Match Trading Group. This is **not** consistent with the core Trade Up hypothesis.

```{r trends, echo=FALSE, warning=FALSE}
CBTU_AVG <- CBTU %>%
  group_by(Scheme) %>%
  summarise_at(
    vars(starts_with("Trading Ratio")),
    list(median = ~ median(., na.rm = TRUE))
  ) %>%
  gather(`Ratio Type`, Value, 2:ncol(.)) %>%
  spread(Scheme, Value)

p <-
  CBTU_AVG %>%
    filter(str_detect(`Ratio Type`, "Trading Ratio Q")) %>%
    ggplot(aes(
      x = `Ratio Type`,
      text = paste0(
        "Control: ", round(Control, 2), "\n",
        "Match Trading: ", round(`Match Trading`, 2)
      )
    )) +
    geom_point(aes(y = Control, colour = "Control"), size = 2) +
    geom_line(aes(y = Control, group = "Control", colour = "Control"), size = 1.25) +
    geom_point(aes(y = `Match Trading`, colour = "Match Trading"), size = 2) +
    geom_line(aes(y = `Match Trading`, group = "Match Trading", colour = "Match Trading"), size = 1.25) +
    scale_x_discrete(labels = c("Q1", "Q2", "Q3", "Q4")) +
    labs(
      title = "Line chart: Quarterly median trading ratios",
      colour = "", x = "", y = "Percentage",
      caption = "(based on data from the School for Social Entrepreneurs)"
    ) +
    ylim(0, 100)
ggplot_ptc(p, "text")

p <-
  CBTU_AVG %>%
    filter(str_detect(`Ratio Type`, "Trading Ratio Change Q")) %>%
    ggplot(aes(x = `Ratio Type`, text = paste0("Control: ", round(Control, 2), "\n", "Match Trading: ", round(`Match Trading`, 2)))) +
    geom_point(aes(y = Control, colour = "Control"), size = 2) +
    geom_line(aes(y = Control, group = "Control", colour = "Control"), size = 1.25) +
    geom_point(aes(y = `Match Trading`, colour = "Match Trading"), size = 2) +
    geom_line(aes(y = `Match Trading`, group = "Match Trading", colour = "Match Trading"), size = 1.25) +
    scale_x_discrete(labels = c("Q1", "Q2", "Q3", "Q4")) +
    labs(
      title = "Line chart: Quarterly increase/decrease in trading ratios",
      colour = "", x = "", y = "Percentage point change"
      ) +
    geom_hline(aes(yintercept = 0)) +
    ylim(-10, 10)
ggplot_ptc(p, "text")

p <-
  CBTU %>%
    select(
      Project, Scheme, Sector, Location,
      contains("Trading Ratio Change Q")
    ) %>%
    gather(Quarter, `Trading Ratio Change`, 5:ncol(.)) %>%
    mutate(Quarter = str_sub(Quarter, -2)) %>%
    ggplot(aes(
      x = Quarter, y = `Trading Ratio Change`, group = Project,
      text = paste0(Project, ": ", round(`Trading Ratio Change`, 2),"\n", Sector)
    )) +
    geom_point(aes(colour = Sector), show.legend = FALSE) +
    geom_line(aes(colour = Sector), show.legend = FALSE) +
    geom_hline(aes(yintercept = 0)) +
    labs(
      title = "Line chart: Quarterly increase/decrease in trading ratios, by group and location",
      x = "", y = "Percentage point change\n"
    ) +
    facet_wrap(vars(Scheme, Location), ncol = 5) +
  theme(legend.position = "none")
ggplot_ptc(p, "text")
```
**Patterns**

This final chart compares the increase/decrease in trading ratios with the "Eligible To Claim Ratio" (ie. the quarterly Trade Up grant as a proportion of total quarterly income). There should be no relationshhip between the two.

```{r patterns, echo=FALSE, warning=FALSE}
p <-
  CBTU %>%
    select(
      Project, Scheme, Location,
      contains("Trading Ratio Change Q"),
      contains("Eligible To Claim Ratio Q")
    ) %>%
    gather(`Ratio Type`, Value, 4:ncol(.)) %>%
    mutate(Quarter = str_sub(`Ratio Type`, -2)) %>%
    mutate(`Ratio Type` = str_sub(`Ratio Type`, 1, -4)) %>%
    replace_na(list(Value = 0)) %>%
    spread(`Ratio Type`, Value) %>%
    ggplot(aes(text = paste0(Location, " - ", Project, ":\nTrading Ratio Change: ", round(`Trading Ratio Change`, 2), "\nEligible To Claim Ratio: ", round(`Eligible To Claim Ratio`, 2)))) +
    geom_point(aes(
      x = `Eligible To Claim Ratio`, y = `Trading Ratio Change`,
      colour = Location
    )) +
    geom_hline(aes(yintercept = 0)) +
    geom_vline(aes(xintercept = 50)) +
    labs(
      title = "Scatter plot: Quarterly increase/decrease in trading ratio, relative to eligible to claim ratio",
      x = "\n\nEligible To Claim Ratio\n\n",
      y = "\n\nPercentage point change\n\n"
    ) +
    facet_grid(Scheme ~ Quarter) +
    xlim(0, 100)
ggplot_ptc(p, "text")
```
