---
title: "2018 Trade Up Analysis (Bayesian)"
author: "Richard Harries"
date: '`r format(Sys.Date(), "%A, %d %B %Y")`'
output: html_document
---
  
```{r import libraries, echo=FALSE, include=FALSE}
library(kableExtra)
library(tidyverse)
library(readxl)
library(tidybayes)
source("Common Tools.R")
```

```{r import data, echo=FALSE, include=FALSE}
CBTU_2017 <-
  read_excel("MT CBTUD PTC 170918.xlsx",
    sheet = "Sheet1",
    skip = 0
  ) %>%
  mutate(Year = 2017) %>%
  select(-c("Sector", "Beneficiary", "Impact Area 1", "Impact Area 2"))

CBTU_2018 <-
  read_excel("CBTU PtC - Roll Out Year Q4 110219.xlsx",
    sheet = "Clean Data",
    skip = 0
  ) %>%
  mutate(Year = 2018) %>%
  select(-c("Sector", "Beneficiary", "Impact Area 1", "Impact Area 2"))

CBTU <- full_join(CBTU_2017, CBTU_2018) %>%
  mutate(
    Scheme = case_when(
      Scheme == "C" ~ "Control",
      Scheme == "MT" ~ "Match Trading"
    )
  )

# Calculate quarterly trading ratios and annual cumulative ratios
# for the baseline and programme years
```

```{r calculate ratios, echo=FALSE, include=FALSE}
for (i in c("Baseline ", "")) {
  for (j in 1:4) {
    CBTU[paste0("Trading Ratio ", i, "Q", j)] <- 100 * CBTU[paste0("Trading Income ", i, "Q", j)] /
      (CBTU[paste0("Trading Income ", i, "Q", j)] + CBTU[paste0("Non-trading Income ", i, "Q", j)])
  }
  CBTU[paste0("Trading Ratio ", i, "Cumulative")] <- 100 *
    rowSums(CBTU[grepl(paste0("Trading Income ", i, "Q"), names(CBTU))], na.rm = FALSE) /
    (rowSums(CBTU[grepl(paste0("Trading Income ", i, "Q"), names(CBTU))], na.rm = FALSE) +
      rowSums(CBTU[grepl(paste0("Non-trading Income ", i, "Q"), names(CBTU))], na.rm = FALSE))
}

for (j in 1:4) {
  CBTU[paste0("Trading Ratio Change Q", j)] <-
    (CBTU[paste0("Trading Ratio Q", j)] - CBTU[paste0("Trading Ratio Baseline Q", j)])
  CBTU[paste0("Eligible To Claim Ratio Q", j)] <-
    (CBTU[paste0("Eligible To Claim Q", j)] / (CBTU[paste0("Trading Income Q", j)] + CBTU[paste0("Non-trading Income Q", j)]))
}

CBTU["Trading Ratio Change Cumulative"] <-
  CBTU["Trading Ratio Cumulative"] -
  CBTU["Trading Ratio Baseline Cumulative"]
```

## Exploratory Data Analysis

Combining the datasets for 2017 and 2018 produces a sample of 128 community businesses, with up to 20 data points recorded for each[^1]. However, the School for Social Entrepreneurs has not yet provided a complete data.

[^1]: Trading income and non-trading income, for each quarter, in the baseline year and programme year, plus four quarterly Trade Up grant awards.

```{r missing data, echo=FALSE, include=TRUE}
CBTU_NA <- CBTU %>%
  group_by(Scheme, Year) %>%
  summarise_at(
    vars(starts_with("Trading Ratio Change Q")),
    list(missing = ~ 100 * sum(is.na(.)) / length(.))
  ) %>%
  gather(`Ratio Type`, Value, 3:ncol(.))

p <- CBTU_NA %>%
#  ggplot() +
  ggplot(aes(
      text = paste0(
        Scheme, " (", Year, "): ", round(Value, 0), "%"
      )
    )) +
  geom_line(aes(x = `Ratio Type`, y=Value,
                group=interaction(Year, Scheme),
                colour=interaction(Year, Scheme)),
            size = 1.25) +
      scale_x_discrete(labels = c("Q1", "Q2", "Q3", "Q4")) +
    labs(
      title = "Line chart: Missing data",
      colour = "", x = "", y = "Percentage",
      caption = "(based on data from the School for Social Entrepreneurs)"
    ) +
    ylim(0, 100)
ggplot_ptc(p, "text")

CBTU <- CBTU %>% drop_na()
```

**Trading ratios**

*Performance*

The average trading ratio in the Match Trading Group appears to *rise* by at least seven percentage points compared with a modest *fall* in Control Group. This is consistent with the core hypothesis of Trade Up, however there is considerable overalap between the two groups and the difference is *not* significant at the 5% level using traditional NHST-style tests.

```{r trading ratio averages, echo=FALSE, include=TRUE}
CBTU %>%
  group_by(Scheme) %>%
  summarise(
    n = n(),
    mean_Change_C = mean(`Trading Ratio Change Cumulative`, na.rm = TRUE),
    median_Change_C = median(`Trading Ratio Change Cumulative`, na.rm = TRUE)
  ) %>%
  kable(
    col.names = c(
      "Group", "n",
      "Mean",
      "Median"
    ),
    digits = 1, format = "html"
  ) %>%
  add_header_above(c(" " = 2, "Percentage point change" = 2)) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "center")

p <- CBTU %>%
  ggplot(aes(Scheme, `Trading Ratio Change Cumulative`,
    text = round(`Trading Ratio Change Cumulative`, 2)
  )) +
  geom_boxplot() +
  geom_point(aes(colour = as_factor(`Year`)), show.legend = TRUE) +
  labs(
    title = "Boxplot: Increase/decrease in trading ratios",
    x = "Trading ratio", y = "Percentage point change",
    colour = "Trade Up\nYear"
  ) +
  coord_flip()
ggplot_ptc(p, "text")

CBTU %>%
  mutate(Group = as.factor(Scheme),
         `Trading Ratio Change` = `Trading Ratio Change Cumulative`) %>%
  kruskal.test(`Trading Ratio Change` ~ Group, .)

CBTU %>%
  mutate(Group = as.factor(Scheme),
         `Trading Ratio Change` = `Trading Ratio Change Cumulative`) %>%
  wilcox.test(`Trading Ratio Change` ~ Group, .)

```

```{r bayes esimates, echo=FALSE, include=FALSE}

graphFileType <- "jpeg"
fileNameRoot <- "TwoGroupTUrobustHet-"
RopeMuDiff <- c(-0.5, 0.5)
RopeSdDiff <- c(-0.5, 0.5)
RopeEff <- c(-0.1, 0.1)

# Load the relevant model into R's working memory:
source("Jags-Ymet-Xnom2grp-MrobustHet.R")

# Generate the MCMC chain:
mcmcCoda <- genMCMC(
  datFrm = as.data.frame(CBTU),
  numSavedSteps = 300000, saveName = "TwoGroupTUrobustHet-",
  yName = "Trading Ratio Change Cumulative", xName = "Scheme"
)

# Get summary statistics of chain:
summaryInfo <- smryMCMC(mcmcCoda,
  RopeMuDiff = RopeMuDiff,
  RopeSdDiff = RopeSdDiff, RopeEff = RopeEff,
  saveName = NULL
)

mcmcMat <- as.matrix(mcmcCoda, chains = TRUE)
mu2 <- mcmcMat[, "mu[2]"]
mu1 <- mcmcMat[, "mu[1]"]
xlim <- range(c(mu1, mu2))
```

```{r bayes output, echo=FALSE, include=TRUE}

plotPost(mu1,
  compVal = 0,
  xlab = bquote(mu[1]), cex.lab = 1.75,
  main = "Control Group", col = "skyblue"
)

plotPost(mu2,
  compVal = 0,
  xlab = bquote(mu[2]), cex.lab = 1.75,
  main = "Match Trading Group", col = "skyblue"
)

plotPost(mu2 - mu1,
  compVal = 0,
  xlab = bquote("muDiff =", mu[2] - mu[1]), cex.lab = 1.75,
  main = "Difference of Means:\nMatch Trading - Control", col = "skyblue"
)

as_tibble(as_tibble(rownames = "id", summaryInfo)) %>%
  select(id, HDIlow, HDIhigh, Mean, Median, Mode, PcntGtCompVal) %>%
  kable(
    col.names = c(
      "Variable", "HDI Lower Bound", "HDI Upper Bound",
      "Mean", "Median", "Mode",
      "Credible %\ngreater than zero"
    ),
    digits = 1, format = "html"
  ) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "center")
```
